---
title: "R Notebook"
output: html_notebook
---

Let's load the datasets of energy consumption by continent and energy consumption by country

```{r}
continent = read.csv("Continent_Consumption_TWH.csv")
country = read.csv("Country_COnsumption_TWH.csv")
```
continent: * The dataset lists up the amount of energy consumption(TWh) group by 7 continents and 3 organizations from year 1990 to 2020. Specifically organizations analyzed in the dataset are OECD, BRICS, and CIS.

country: * It is the dataset that lists up the amount of power consumption of 44 countries from year 1990 to 2020.
```{r}
str(continent)
```
```{r}
str(country)
```
install the necessary packages


```{r}
library(tidyverse)
library(ggplot2)
library(rmarkdown)
library(knitr)
library(ggthemes)

```
let us look at the dataset a little bit, lets look at the years the data has been archived
```{r}
unique(continent$Year)
```
lets see if we have any missing data
```{r}
sum(is.na(continent))
```
```{r}
head(continent, 5)
```
```{r}
head(country, 5)
```
```{r}
summary(continent)
```
Let us visualise the data
lets start with how energy consumtion changed since 1990 to 2020
```{r}
ggplot(
    filter(gather(continent, continent, Twh, World:CIS), continent == "World"), 
    aes(Year, Twh))+
    geom_point()+
    geom_line()
```
```{r}
filter(gather(continent,continent,Twh,World:CIS),continent == "World") %>% 
    mutate(percent_increase = (Twh - lag(Twh, default = first(Twh)))/ lag(Twh, default =first(Twh))) %>%
    ggplot(aes(Year, percent_increase))+
    geom_point() +
    geom_smooth(method = "lm")
```
```{r}
ggplot(
    filter(gather(continent,continent,Twh,World:CIS),continent != "World"& continent != "CIS" & continent != "BRICS" & continent !=  "OECD"),
    aes(Year,Twh, color = continent)) +
    geom_point() + 
    geom_smooth()
```
```{r}
ggplot(
    filter(gather(continent,continent,Twh,World:CIS),continent != "World"& continent != "CIS" & continent != "BRICS" & continent !=  "OECD" & continent !=  "Asia"),
    aes(Year,Twh, color = continent)) +
    geom_point() + 
    geom_smooth() +
    facet_wrap(~continent)
```
```{r}
ggplot(
    filter(gather(continent,continent,Twh,World:CIS),continent != "World"& continent != "CIS" & continent != "BRICS" & continent !=  "OECD",
           continent != "Asia",continent != "Africa",continent != "Latin.America",continent != "Middle.East",continent != "Pacific"),
    aes(Year,Twh, color = continent)) +
    geom_point() + 
    geom_smooth() +
    facet_wrap(~continent)
```
```{r}
ggplot(
    filter(gather(continent,continent,Twh,World:CIS),continent != "Africa" & continent != "Asia" 
           & continent !=  "Europe" & continent !=  "Latin.America" 
           & continent !=  "Middle.East" & continent !=  "North.America" 
           & continent !=  "Pacific"),
    aes(continent,Twh, fill = continent)) +
    geom_col()
```
```{r}
ggplot(
    filter(gather(continent,continent,Twh,World:CIS),continent != "Africa" & continent != "Asia" 
           & continent !=  "Europe" & continent !=  "Latin.America" 
           & continent !=  "Middle.East" & continent !=  "North.America" 
           & continent !=  "Pacific" & continent !=  "World"),
    aes(Year,Twh, color = continent)) +
    geom_point() + 
    geom_line()
```
```{r}
gather(select(continent, "Year", "BRICS", "CIS", "OECD"), economic_alliances, Twh, BRICS:OECD) %>%
    ggplot(aes(Year, Twh, color = economic_alliances)) +
    geom_point() +
    geom_smooth(method = "lm")
```
```{r}
gather(select(continent,"Year","BRICS", "CIS", "OECD"),economic_alliances, Twh, BRICS:OECD) %>% 
    ggplot(aes(economic_alliances, Twh, fill = economic_alliances))+ geom_col()
```
```{r}
gather(select(continent,"Year","BRICS", "CIS", "OECD"),economic_alliances, Twh, BRICS:OECD) %>% 
    ggplot(aes( Twh, fill = economic_alliances))+ geom_boxplot()
```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
